Evidence of Magnetic Saturation from Stellar Flare Rates

### 2023-10-30 src/tex/ms.tex
young planet host flare rate comparison

% Title
\title{Refined Flare Rates for Stars Younger than 250\,Myr}

@@

We followed the machine learning flare-identification methods presented in
\cite{feinstein20a}. This method relies on the fact that all flare events
have similar time-dependent morphologies generally described as a sharp rise
and an exponential decay. This method uses a convolutional neural network (CNN),
\texttt{stella} \citep{feinstein20b}, trained on by-eye validated flare events from
TESS Sectors 1 and 2 to identify flare events in TESS 2-minute data
\citep{guenther19_flares, feinstein20a}. The benefits of the CNN is that it is
insensitive to the stellar baseline, since it is trained to look only for the
flare morphology. This means that the peaks of rotational modulation driven by
stellar heterogeneities, which is readily seen in the light curves of young stars,
are not accidentally identified as flares. In this way, we are able to
build a sample of flares which is unbiased towards low-amplitude/low-energy flares,
which are often not identified in traditional sigma-outlier identification
methods \citep{chang15}.

@@

\subsection{Modeling Flare Properties}

@@

From these fits, we performed a series of quality checks to ensure the flares
in our sample are real. First, we removed any flares with best-fit FWHM $ < 0$
and $A < 0$. Second, we removed any flares with a large error on the amplitude,
$\sigma_A > 10$ and flares where $\sigma_A > A$. Third, we removed flares where
$A < 2 \times \textrm{RMS}$, where the RMS is calculated from the 10~hour baseline
around the best-fit $t_\textrm{peak}$. These cuts are indicative of a poorly fit
flare model due to the flare being overwhelmed by noise in the surrounding baseline
light curve. We choose to include flares with a probability $P \geq 75\%$ of being
a true flare. After these checks, we have a flare sample of 29,128~flares originating
from 3,983 stars (Figure~\ref{fig:flare_distribution}). 83.6\% of the flares have
probabilities of $P \geq 90\%$ of being true; 64.9\% of the flares have probabilities
of $P \geq 99\%$ of being true.

@@

@@

Additionally, we analyzed the TESS light curves for all of the ULLYSES targets, when
available at 2-minute cadence. We identified a total of \textcolor{red}{n} flares.
This is inconsistent with the overall flare statistics identified in this work, and is
overall unexpected for young stars which are known to be more magnetically active
\citep{ilin20, feinstein20b}. This could suggest that there is some bias in the way
the ULLYSES sample was constructed, and that it is unexpected to find any flares in the
FUV data. Alternatively, it could be that the ongoing accretion luminosity results
in a brighter host star, such that the flares we normally would see are drowned
out by the increased baseline luminosity. We do not include this flare in our
overall statistical analysis of TESS flare rates for young stars.

@@

\end{document}

>>>>

% Title
\title{Evidence of Magnetic Saturation from Stellar Flare Rates}

@@

We followed the machine learning flare-identification methods presented in
\cite{feinstein20a}. This method relies on the fact that all flare events
have similar time-dependent morphologies generally described as a sharp rise
and an exponential decay. This method uses a convolutional neural network (CNN),
\texttt{stella} \citep{feinstein20}, trained on by-eye validated flare events from
TESS Sectors 1 and 2 to identify flare events in TESS 2-minute data
\citep{guenther19_flares, feinstein20a}. The benefits of the CNN is that it is
insensitive to the stellar baseline, since it is trained to look only for the
flare morphology. This means that the peaks of rotational modulation driven by
stellar heterogeneities, which is readily seen in the light curves of young stars,
are not accidentally identified as flares. In this way, we are able to
build a sample of flares which is unbiased towards low-amplitude/low-energy flares,
which are often not identified in traditional sigma-outlier identification
methods \citep{chang15}.

@@

\subsubsection{Modeling Flare Properties}

@@

\subsubsection{Flare Quality Checks}

The \texttt{stella} CNNs were initially trained on data from TESS Sectors 1 and 2.
Because the noise properties from sector to sector varies, the CNNs are unable to accurately
capture all of these changes. Additionally, the CNNs were trained on a sample of \textcolor{red}{n}
stars, which does not necessarily encapsulate all types of variable stars or a sufficient distribution
of variable stars (e.g., eclipsing binaries and RR Lyraes). As such, we apply the additional
quality checks presented in \cite{Feinstein22} to ensure our flare sample has little to no contamination
from other sources. We remove flares which satisfy the one or more of the following criteria:

\begin{enumerate}
  \item The amplitude of the flare must be $> 0.01$, or 1\%. This follows the limits set by
        \cite{feinstein20}.
  \item The amplitude of the flare (A) must be greater than $1.5\times$CDPP \citep[Combined Differential
        Photometric Precision;][]{christiansen12}, similar to \cite{Feinstein22}.
  \item For flares originating from stars with $T_\textrm{eff} > 5000$, the amplitude of the flare
        must be greater than $12\times$CDPP. This is because the most massive stars tended to have
        much larger scatter in the TESS noise.
  \item The fitted equivalent duration (ED) must be positive $\textrm{ED} > 0$.
  \item The fitted full-width half-maximum (FHWM) and A must be $> 0$.
  \item The error on the A must be $\sigma_A < A$. This removed poorly constrained flares and often
        flares which were buried in the TESS noise.
\end{enumerate}

We choose to include flares with a probability $P \geq 75\%$ of being
a true flare. After these checks, we have a flare sample of 28,822~flares originating
from 3,979 stars (Figure~\ref{fig:flare_distribution}). 83.8\% of the flares have
probabilities of $P \geq 90\%$ of being true; 65.0\% of the flares have probabilities
of $P \geq 99\%$ of being true.

@@

\subsection{Correlation to Other Metrics of Magnetic Saturation}

\subsection{Comparison to Young Planet Host Stars}

The all-sky observing strategy of TESS has revealed a new population of young transiting exoplanets.
Characterizing the local environment for these planets is crucial to understanding their subsequent evolution
to forming the mature population of planets. It is debated whether stellar flares are beneficial or detrimental
to exoplanets. On one hand, \textcolor{red}{creates life}. On the other hand, stellar flares and affiliated
coronal mass ejections can permanently alter atmospheric compositions \citep{chen21} and increase the amount of
atmospheric mass stripped during the early stages of planet evolution \citep{feinstein20}. To understand the
evolution of these new insightful young transiting exoplanets, we compare their measured flare rates to a more
statistical sample of stars with similar ages and $T_\textrm{eff}$.

We measured the flare rates of planet hosting stars $< 300$\,Myr, comparable to the ages of our primary sample.
We followed the methods outlined in Section~\ref{sec:methods} to detect and vet flares. For each star, we created
an equivalent sample with respect to both age and $T_\textrm{eff}$ to compare the flare rates too. We considered
stars with ages $\pm 30$\,Myr of the planet hosting star and $T_\textrm{eff} \pm 1000$\,K. We calculated the overall
flare rate as Equation~\ref{eq:fr}. We report the flare rates of host star and the comparable sample in Table~\ref{tab:yprates}.
For the comparable sample, we report the median flare rate, and the lower 16\textsuperscript{th} and upper 84\textsuperscript{th}
percentiles.

\begin{figure}[ht!]
    \script{yp_rates.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/yp_rates.pdf}
        \caption{
            Comparison of flare rates from young planet host stars with respect to
            a comparable sample with respect to age [Myr] and $T_\textrm{eff}$ [K].
        }
        \label{fig:yp_rates}
    \end{centering}
\end{figure}

\begin{deluxetable}{l r r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{Young Planet Host Flare Rates\label{tab:yprates}}
\tablehead{
\colhead{Host Name} & \colhead{Age} & \colhead{Flare Rate} & \colhead{Comp. Sample} \\
 & \colhead{[Myr]} & \colhead{[day$^{-1}$]} & \colhead{Flare Rate [day$^{-1}$]}
}
\startdata
TOI 1227 &  $11 \pm 2$ &  0.008  & $0.119_{-0.046}^{+0.148}$ \\
HIP 67522 & $17 \pm 2$ &  0.169  & $0.046_{-0.020}^{+0.069}$ \\
AU Mic & $22 \pm 3$ &  2.218  & $0.133_{-0.046}^{+0.171}$ \\
V1298 Tau & $23 \pm 4$ &  0.022  & $0.074_{-0.033}^{+0.091}$ \\
HD 109833 & $27 \pm 3$ & 0.000  & $0.046_{-0.020}^{+0.070}$ \\
KOI-7913 & $36 \pm 10$ &  0.031  & $0.134_{-0.045}^{+0.162}$ \\
KOI-7368 & $36 \pm 10$ &  0.029  & $0.064_{-0.030}^{+0.087}$ \\
DS Tuc & $45 \pm 4$ &  0.420  & $0.056_{-0.024}^{+0.078}$ \\
TOI 942 & $50_{-20}^{+30} $ &  0.040  & $0.086_{-0.038}^{+0.117}$ \\
TOI 451 & $120 \pm 10$ &  0.128  & $0.071_{-0.029}^{+0.085}$ \\
HIP 94235 & $133_{-20}^{+15}$ &  0.020  & $0.051_{-0.017}^{+0.079}$ \\
TOI 1860 & $133 \pm 26$ &  0.008  & $0.067_{-0.025}^{+0.086}$ \\
TOI 1807 & $180 ± 40$ &  0.013  & $0.060_{-0.022}^{+0.085}$ \\
HD 18599 & $200_{-70}^{+200} $ & 0.000 & $0.025_{-0.015}^{+0.044}$ \\
TOI 2076 & $204 \pm 50$ & 0.000 & $0.092_{-0.045}^{+0.091}$ \\
HD 110082 & $250_{-70}^{+50} $ & 0.000 & $0.047_{-0.020}^{+0.067}$ \\
\enddata
\end{deluxetable}

@@

Additionally, we analyzed the TESS light curves for all of the ULLYSES targets, when
available at 2-minute cadence. We identified a total of \textcolor{red}{n} flares.
This is inconsistent with the overall flare statistics identified in this work, and is
overall unexpected for young stars which are known to be more magnetically active
\citep{ilin20, feinstein20}. This could suggest that there is some bias in the way
the ULLYSES sample was constructed, and that it is unexpected to find any flares in the
FUV data. Alternatively, it could be that the ongoing accretion luminosity results
in a brighter host star, such that the flares we normally would see are drowned
out by the increased baseline luminosity. We do not include this flare in our
overall statistical analysis of TESS flare rates for young stars.

@@

\end{document}
feinstein20
